---
title: "Non-Parametric Tests for Location and Dispersion"
author: "Kamesh Dubey"
date: "`r Sys.Date()`"
output: html_document
---


# **1. Test For Location**

## **1.1 Sign Test**

---

The one-sample sign test is used to determine if the median of a single sample $\tilde{x}$ is equal to a specified value. 


**Step 1:** Define the Hypothesis

\[ 
H_0: \tilde{x} = M_0 
\]
\[ 
H_1: \tilde{x} \neq M_0 
\]

**Step 2:** Define:

- \(N^+\) =  **# $x_i$ > \(M_0\).**
- \(N^-\) = **# $x_i$ < \(M_0\).**

Using above, count the number of observations above (\(N^+\)) and below (\(N^-\)) the hypothesized median \(M_0\)

<br>

**Step 3:** The test statistic is the smaller of \(N^+\) and \(N^-\), which follows a binomial distribution \( \text{Binomial}(N, 0.5) \), where \( N = N^+ + N^- \). 

<br>

**Step 4:** The p-value from this binomial test, comparing \(N^+\) and \(N^-\) to the expected distribution under \(H_0\), determines if we reject \(H_0\).

```{r}
# Sample data
x = c(2, 3, -1, 4, 5, -2, 3, -3, 6, 7, 1, -1)

                     # step1 
#Hypothesized median
median = 0

                    #step 2 
#Calculate the differences from the hypothesized median
differences = x - median

# Count the number of positive, negative, and zero differences
positive = sum(differences > 0)
negative = sum(differences < 0)
zero = sum(differences == 0)


# Perform the sign test using the binom.test function
# Note: We exclude zero differences from the test
result = binom.test(x = c(positive, negative), alternative = "two.sided")

print(result)

```
<br>

**Interpreting the results**

The p-value of 0.3877 suggests that there is insufficient evidence to reject the null hypothesis at a significance level of 0.05. Therefore, based on this test, we cannot conclude that the median of the sample significantly differs from the 0.

<br>
<br>

#### **Lets say our data has some 0's as observation, i.e. some observations are equal to mean. What to do in that case ?**
```{r}
y = c(2, 3, 5, 0, 4, 5, 1, -2, 3, -3, 6, 7, 1, -1)

# Hypothesized median
median0 = 0

# Calculate the differences from the hypothesized median
differences = y - median0



# Count the number of positive, negative, and zero differences
positive <- sum(differences > 0)
negative <- sum(differences < 0)
zero <- sum(differences == 0)


result = binom.test(x = c(positive, negative), alternative = "greater")

print(result)
```
<br>

**Interpreting the results**

The p-value of 0.04614 suggests that the *observed data provides evidence against the null hypothesis*. Hence, We reject Null in favour of alternate.

**Note:** In the context of the sign test, \( x_i = 0 \) observations are disregarded because they provide no evidence for or against the alternative hypothesis. Including them would not affect the test's outcome, as the sign test focuses solely on the direction of differences rather than their magnitude. Thus, excluding \( x_i = 0 \) simplifies the analysis without compromising its validity.

<br>
<br>
<br>


#### **What happens when number of observation in the sample are large ?**

When the number of observations is large, such as $|ð‘‹|$ > 30, we can still perform a one-sample sign test, but the **exact binomial test** may not be the most appropriate approach due to computational limitations and because of that the normal approximation might be more accurate in this case.

```{r}
y = runif(40, min = -10, max = 10)


# Hypothesized median
median0 = 0

# Calculate the differences from the hypothesized median
differences = y - median0



# Count the number of positive, negative, and zero differences
positive <- sum(differences > 0)
negative <- sum(differences < 0)
zero <- sum(differences == 0)

# Total number of observations
n <- length(y)

# Calculate the standard error
se <- sqrt(n * 0.25) #sqrt(np(1-p))

# Calculate the z-score note N+ is used as test statistics
z <- (positive + 0.5 - 0.5 * n) / se

# Calculate the p-value using the normal approximation
p_value <- 2 * pnorm(abs(z), lower.tail = FALSE)

# Output the result
cat("Z-score:", z, "\n")
cat("P-value:", p_value, "\n")

```
Here, 
      p>0.05. Then, we can't reject null. Hence, we can reject null hypothesis, ie $\tilde{x}=0$


