---
title: '#Hypothesis Testing: Simple Linear Model'
author: "Kamesh Dubey"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## 1. Generate Data

--- 

```{r}
set.seed(0)

age = round(runif(20, min=20, max = 80))

heartrate = (220-age)
heartrate = round(heartrate + rnorm(20, mean=0, sd=15))

cat("Age:", age )
cat("\n\nHearbeat:", heartrate)
```
## Create a Dataframe

---

```{r}
data = data.frame(Age=age, Heartrate=heartrate)

print(data)
```

## Plot Age and Heartbeart

---

```{r}
par(plt=c(0.2, 0.8, 0.2, 0.75))
plot(age, heartrate,
     xlab = "Age", ylab="Heart Rate",
     main = "Age Vs. Heart Rate scatterplot",
     pch=16, col="steelblue")
```

## Simple Linear Model

---

```{r}
model = lm(heartrate~age)
summary(model)
```

#### Observations

- using p-values, we can see that both Intercept and age are significant.

- R-Squared: 0.743, Adjusted R-Squared:0.729. We can conclude that age explain the heart rate quite well

- The equation for heart rate from **Linear Model** can be expressed as:$ \text{heart rate} = 219.1365 - 1.0221 \times \text{age}$




## Residual Analysis

---

```{r}
res = residuals(model)
pred = fitted(model)

cat("Fitted-Values: ", pred)
cat("\n\nResidual: ", res)
cat("\n\nSummary of Residual\n")
print(summary(res))
```
- From Residual summary we have: 
                              $\sum_{i=1}^{20} e_{i}=0$ or $$\bar{e_{i}}=0$$



```{r echo=FALSE}
par(plt=c(0.2, 0.8, 0.2, 0.75))

plot(pred, res,
     xlab="Fitted-Values", ylab="Residuals",
     main="Residual Vs. Fitted Values",
     pch=16, col="steelblue")
abline(h=0, col="red", lwd=2, lty=2)

```

#### From the Residual vs Fitted Values plot, We can conclude that 
- Residuals are Randomly Scattered around Zero mean.
- No Systematic pattern can be seen in the plot.
- Variance of Residual seems Constant
- No sign of outliers.


## Testing of Hypothesis

---

```{r}
n = length(age)

beta_0_hat = coef(model)[["age"]]

RSS=sum(res^2)

sigma_hat = sqrt(RSS/n-2)

Sxx = sqrt(sum(age=mean(age)^2))
Std_err_beta_0 = sigma_hat/Sxx

cat("beta_0=", beta_0_hat)
cat("\n\nSigma_hat=", sigma_hat)
cat("\n\nSE(beta_0)=", Std_err_beta_0)
```

#### Test for $\betas$**

---

- **for $\beta_{0}$**

Consider

$$H_0: \beta_{0} = 0 \quad \text{vs} \quad H_1: \beta_{0} \neq 0 $$
Then,
```{r}
t_0 = (beta_0_hat-(-0))/Std_err_beta_0

cat("       Under Ho,\n                   t-statistic value= ", t_0)

p_value = pt(t_0, 20, lower.tail=FALSE)

cat("\n\n     The corresponding p-value for the test statistic will be:\n\n                     p-value=", p_value)
```
The p-value of 0.99 (>>0.05) tell us that $\beta_{0}=0$ is not likely. 

---

This time, Let's Consider 


$$H_0: \beta_{0} = -1 \quad \text{vs}  \quad H_1: \beta_{0} \neq -1$$

```{r}
Sxx = sqrt(sum(age=mean(age)^2))
Std_err_beta_0 = sigma_hat/Sxx

t = (beta_0_hat-(-1))/Std_err_beta_0
cat("Under Ho,\n            t-statistic value= ", t )

p_value = pt(t, 20, lower.tail=FALSE)

cat("\n\n The corresponding p-value for the test statistic will be:\n\n              p-value=", p_value)
```
The p-value of 0.99 (>>0.05) tell us that $\beta_{0}=-1$ is also not likely. 

